Board Game Recommendation

### 2015-03-05 paper.Rnw
Revisions and Additions
Revised the modeling section, added some data for some tables

\newcommand{\hh}[1]{{\color{magenta} #1}}

@@

\section{Introduction}

@@

There are two types of particular information that we will be interested in: user ratings of the games and characteristics about the game (referred to as classifiers in the modeling section). The characteristics we're interested in fall into these categories as outlined in table~\ref{tab:categories}

@@

The ratings for each game are on a scale from 1 to 10, with a score of 1 indicating a really bad game and a score of 10 being the best possible. Each of the ratings are paired with a unique user identification number as well as a game number. Combined with the classifiers data, we can provide very specific ratings for each individual user.

\hh{Give an example of how the data that you have scraped looks like. This would be a good point to introduce a bit of the notation that you are using below, like the $y_{u,i}$. }

\hh{Aren't you continuously scraping? You should mention this at this point, and also talk a bit about how you have implemented the scraping.}

@@

Note for editing: I know this model is a mess at the moment, just getting the general ideas down on paper. 
\hh{this is not nearly as messy as you make it out to be, but you need to make sure to define every little piece of notation as you, because otherwise your reader will be lost.}
\subsubsection{Global Effects}
 The general idea is to remove the global mean from all ratings to get a residual for each observation. We'll group these residuals by game, and then find an effect for that game. Subtracting this effect from the residual, we'll take those residuals and repeat the same for each user. The motivations for using the formulas that follow can be found in Scalable Collaborative Filtering with Jointly Derived Neighborhood Interpolation Weights (make this a bib reference, \hh{you still need to paraphrase the motivation here.}).

As we proceed $y_{u,i}$ will denote the rating by user $u$ for game $i$ \hh{what scale is the rating on?}. Then, we have $$y_{u,i} = \mu_{global} + \mu_{i} + \mu_{u} + error$$

However, we'll approach this in steps, as follows: \hh{what is the benefit of using a hierarchical approach?}

\begin{center}
\begin{align}
y_{u,i} &= \mu_{global} + error_{global} \\
error_{global} &= \mu_{i} + error_{game} \\
error_{game} &= \mu_{u} + error_{user} 
\end{align}
\end{center}

And we'll use the estimation method outlined in the paper (need ref) for each, as follows below:

\begin{center}
\begin{align}
\mu_{global} = \frac{\sum_{u = 1}^{N_{u}}\sum_{i = 1}^{N_{games}}}{1}
\end{align}
\end{center}
\hh{something went wrong in the formula - what are you summing over?}

>>>>

\newcommand{\hh}[1]{{\color{magenta} #1}}
\newcommand{\sam}[1]{{\color{blue} #1}}

@@

<<loader, echo = FALSE>>=
library(xtable)
load("paper_stuff")
@
\section{Introduction}

@@

There are two types of particular information that we will be interested in: user ratings of the games and characteristics about the game (referred to as classifiers in the modeling section). The characteristics we're interested in fall into these categories as outlined in table~%\ref{tab:categories}

@@

<<classifiers, echo = FALSE, results = 'asis'>>=
xtable(paper_stuff$classifiers, caption = c("Classifiers Data, as seen in Database", "Classifiers Table"))
@

The ratings for each game are on a scale from 1 to 10, with a score of 1 indicating a really bad game and a score of 10 being the best possible. Each of the ratings are paired with a unique user identification number as well as a game number. Combined with the classifiers data, we can provide very specific ratings for each individual user.
<<ratings, echo = FALSE, results = 'asis'>>=
xtable(paper_stuff$ratings, caption = c("Ratings Data, as seen in Database", "Ratings Table"))
@

To keep the information as compact as possible, individual players and games are referred to by an ID number, which corresponds to a row in one of the key tables. These key tables contain information about the players and game that we don't necessarily need to include with each rating or classifier. This greatly reduces the size of the database needed to store and work with the information.
<<player_key, echo = FALSE, results = 'asis'>>=
xtable(paper_stuff$player_key, caption = c("The Player Key relates important info about the individual to the userid", "Player Key Table"))
@

<<game_key, echo = FALSE, results = 'asis'>>=
xtable(paper_stuff$game_key[,1:6], caption = c("The Game Key gives useful information about the game", "Game Key Table"))
@

Not all of the data was scraped before creating the applicaiton. Data for 6,088 games was scraped and stored across the four tables. The reason for not scraping all of the data is twofold. First of all, for those 6,088 games there are over 3 million unique ratings. If we were to expand it to all 75,756 games known to BoardGameGeek, then our data set would be crippilingly large for the small desktop I have to work on. Secondly, this will force us to implement a continuous data scraping procedure. Each time the person uses the app, all we have to do is make sure we have all of their current information correct. We look through our ratings and can check them with their userpage on BoardGameGeek. From this, we can revise old ratings or add new ones, as well as check to see that all of the games that they've rated are in our database!

\hh{Give an example of how the data that you have scraped looks like. This would be a good point to introduce a bit of the notation that you are using below, like the $y_{u,i}$. }
\sam{I still need to tie this back to the modeling section.}

@@

Note for editing: I know this model is a mess at the moment, just getting the general ideas down on paper. 
\hh{this is not nearly as messy as you make it out to be, but you need to make sure to define every little piece of notation as you, because otherwise your reader will be lost.}
\subsubsection{Global Effects}
The general idea behind global effects is to account for the natural biases present in particular games or players. For example, we might have two players who both play and enjoy the same game the same amount. However, one person might give this game a score of 10 while the other gives it a score of 9. Despite their overall enjoyment being the same, perhaps the latter player just tends to be a harsher rater in general. The global effect for this person would then be lower than that of the much more cheery fellow who gave the game a 10. Similarly, this effect can be present for games due to any number of reasons.

We'll estimate these effects one at a time using a heirarchical structure. After estimating the effect, we'll be left with a residual, which will be used as the response for the next level.
The global effects formual is

\begin{equation}
r_{u,i} = \theta_u x_{u,i} + error
\end{equation}

Where for the first step, $r_{u,i}$ will be the rating from user $u$ for game $i$ and $x_{u,i}$ will be an indicator variable that is 1 if player $u$ has rated game $i$ and 0 otherwise. In the second step $r_{u,i}$ will be the residual from the first portion, the subscript on $\theta$ becomes $i$ and $x_{u,i}$ will be an indicator variable that is 1 if the rating is for game $i$ and 0 otherwise.

To get an estimate of $\theta$ at either step, we'll use the average of the nonzero values which are shunk by the fitting parameter $\alpha$ to avoid overfitting. The sum is over either all ratings for user $u$ or all ratings for game $i$, depending on whether you're estimating $\theta_u$ or $\theta_i$. Either way, the structure is the same.
\begin{equation}
\hat{\theta_u} = \frac{\sum r_{u,i}}{n_u + \alpha}
\end{equation}

%  The general idea is to remove the global mean from all ratings to get a residual for each observation. We'll group these residuals by game, and then find an effect for that game. Subtracting this effect from the residual, we'll take those residuals and repeat the same for each user. The motivations for using the formulas that follow can be found in Scalable Collaborative Filtering with Jointly Derived Neighborhood Interpolation Weights (make this a bib reference, \hh{you still need to paraphrase the motivation here.}).
% 
% To begin $r_{u,i}$ will denote the rating by user $u$ for game $i$, where $y_{u,i}$ is a number between 1 and 10. Then, we have $$y_{u,i} = \mu_{global} + \mu_{i} + \mu_{u} + error$$
% 
% However, we'll approach this in steps, as follows: \hh{what is the benefit of using a hierarchical approach?}
% 
% \begin{center}
% \begin{align}
% y_{u,i} &= \mu_{global} + error_{global} \\
% error_{global} &= \mu_{i} + error_{game} \\
% error_{game} &= \mu_{u} + error_{user} 
% \end{align}
% \end{center}
% 
% And we'll use the estimation method outlined in the paper (need ref) for each, as follows below:
% 
% \begin{center}
% \begin{align}
% \mu_{global} = \frac{\sum_{u = 1}^{N_{u}}\sum_{i = 1}^{N_{games}}}{1}
% \end{align}
% \end{center}
% \hh{something went wrong in the formula - what are you summing over?}
